---
title: "Eating Out"
author: "Chris Gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## There were two years in which the General Social Survey asked about the frequency of eating out.
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(haven)
require(knitr)
require(foreign)
```
```{r, eval=TRUE}
GSS <-read_dta(file="GSS1998.dta")
GSS <- readstata13::read.dta13(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```




#### Plot barplots tigerstats way
```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```

According to the statistical data represented in the bar plot from thetiger stat, approximately 80% of individuals identifying as white reported dining out. In contrast, around 68% of the black population surveyed indicated a preference for eating out, while approximately 72% of individuals from other ethnic backgrounds expressed the same behavior. These findings suggest a higher proportion of white individuals engaging in dining out activities compared to the other two demographic groups under consideration.




#### Plot barplots ggplot way

```{r}
dd2 <- mydata %>% group_by(race,eatout) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=eatout))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=eatout)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Utilizing an alternative visualization technique, specifically a box plot, further validates the initial observations that white individuals exhibit a higher propensity for dining out. The data suggests that out of the surveyed white population, 80% engage in eating out activities. By representing this information using a box plot, we can reaffirm the dominant trend of white individuals' higher preference for dining out compared to other two groups.


### Numerical Summaries   

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```

Upon analyzing the numerical summary data, it is evident that white individuals have a significantly higher proportion of dining out, with approximately 83% of the surveyed population engaging in this behavior. In comparison, around 69% of black individuals and 72% of individuals from other ethnic backgrounds reported participating in eating out activities. These findings strongly support the initial conclusion that white individuals exhibit a higher frequency of dining out compared to the other demographic groups surveyed.


## Inferential  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The statistical analysis demonstrates a remarkably low p-value of 0.000002287, which indicates a highly significant difference in the dining-out behavior between white individuals and the other two groups.

## Fisher

```{r}
fisherXY(mydata$race,mydata$eatout)
```
The analysis reveals a statistically significant difference in the dining-out behavior between white individuals and the other two groups. With a p-value of 0.00001183, the odds of a white person eating out are 2.2 times higher compared to a black individual. Additionally, the p-value of 0.00721 indicates that the odds of white individuals eating out are 1.9 times higher compared to individuals from other ethnic backgrounds. These findings provide robust evidence supporting the conclusion that white individuals have a significantly higher likelihood of engaging in dining-out activities compared to both black individuals and individuals from other ethnic backgrounds.

```{r}
fisherXY(mydata$race,mydata$eatout,printAll = TRUE)
```
Upon further examination of the additional data, it indicates that black individuals have approximately 0.86% higher odds of eating out compared to individuals from the other category. The p-value of 0.5993 between the black and other category suggests that this difference is not statistically significant. In other words, the observed disparity in dining-out behavior between black individuals and those from the other category may occur due to random chance rather than a meaningful distinction.


##Conclusion 


